---
title: "Exercises"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski, Lars Viklund"
date: "Feb., 2021"
output: html_document
---


## 1.

In the previous module about Profiling we wrote a function for computing the sum of the elements of each column:

```{r eval=TRUE}
library(tictoc)
A <- matrix(1.0, 5000, 5000)
sumcol <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  colsm <- rep(0,l) #create a vector to save the sums
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:l){
      s <- s + B[i,j]
    }
    colsm[j] <- s
  }
  return(colsm)
}

tic()
res1 <- sumcol(A)
toc()
```

The bottleneck in this function was the for loop. In this exercise, we will vectorize the heavy loop.
Replace the *FIXME* strings with the corresponding R code:

```{r eval=TRUE}
library(tictoc)
A <- matrix(1.0, 5000, 5000)
sumcol_vec <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  colsm <- rep(0,l) #create a vector to save the sums
  
  for(j in 1:l){  
    colsm[j] <- sum(B[,j])
  }
  return(colsm)
}

tic()
res2 <- sumcol_vec(A)
toc()
```

# 2.

The function *sdcol* which computes the standard deviation of each column of a matrix A of size 10000x3 using for loops. This function has a strong C accent. 

Compare the performance of this function with the counterpart *apply()* function for computing the standard deviation. 

Finally, compare the performance of these two functions with the colSds() function from the *library(matrixStats)* package.


```{r eval=TRUE}
A <- matrix(sample(0:1, 1000000, replace = TRUE),1000000,3)

sdcol <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  n <- nrow(B)  #obtain the number of rows
  colsm <- rep(0,l) #create a vector to save the means
  colsd <- rep(0,l) #create a vector to save the sd
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:n){
      s <- s + B[i,j]
    }
    colsm[j] <- s/n
  }
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:n){
      s <- s + (B[i,j] - colsm[j])^2
    }
    colsd[j] <- sqrt( s/(n-1))
  }
  
  return(colsd)
}
library(tictoc)
tic()
res1 <- sdcol(A)
toc()

tic()
res2 <- apply(A,2,sd)
toc()

#install.packages("matrixStats")
library(matrixStats)
tic()
res3 <- colSds(A)
toc()
```

What function performed best? The function with best performance was *colSds()*
What could be the difference the other two? This is a compiled function besides that it avoids
setting environment variables and duplication of data.

Main message: if you have a compiled function available use it instead of writing your own code.



# 3.

Use the *Rcpp* package to write a C++ function that adds three float numbers:

```{r eval=TRUE}
library(Rcpp)

cppFunction('float sum_cpp(float a, float b, float c ) {
  float sumc = a + b + c;
  return sumc;
}')
sum_cpp
sum_cpp(2,4,6)
```
