---
title: "Exercises"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski, Lars Viklund"
date: "Feb., 2021"
output: html_document
---


## 1.

In the previous module about Profiling we wrote a function for computing the sum of the elements of each column:

```{r eval=TRUE}
library(tictoc)
A <- matrix(1.0, 5000, 5000)
sumcol <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  colsm <- rep(0,l) #create a vector to save the sums
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:l){
      s <- s + B[i,j]
    }
    colsm[j] <- s
  }
  return(colsm)
}

tic()
res1 <- sumcol(A)
toc()
```

The bottleneck in this function was the for loop. In this exercise, we will vectorize the heavy loop.
Replace the *FIXME* strings with the corresponding R code:

```{r eval=TRUE}
library(tictoc)
A <- matrix(1.0, 5000, 5000)
sumcol_vec <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  colsm <- rep(0,l) #create a vector to save the sums
  
  for(j in 1:l){  
    colsm[j] <- sum(B[,j])
  }
  return(colsm)
}

tic()
res2 <- sumcol_vec(A)
toc()
```

# 2.

The function *sdcol* which computes the standard deviation of each column of a matrix A of size 10000x3 using for loops. This function has a strong C accent. 

Compare the performance of this function with the counterpart *apply()* function for computing the standard deviation. 

Finally, compare the performance of these two functions with the colSds() function from the *library(matrixStats)* package.


```{r eval=TRUE}
A <- matrix(sample(0:1, 1000000, replace = TRUE),1000000,3)

sdcol <- function(B) {
  l <- ncol(B)  #obtain the number of columns
  n <- nrow(B)  #obtain the number of rows
  colsm <- rep(0,l) #create a vector to save the means
  colsd <- rep(0,l) #create a vector to save the sd
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:n){
      s <- s + B[i,j]
    }
    colsm[j] <- s/n
  }
  
  for(j in 1:l){  
    s <- 0
    for(i in 1:n){
      s <- s + (B[i,j] - colsm[j])^2
    }
    colsd[j] <- sqrt( s/(n-1))
  }
  
  return(colsd)
}
library(tictoc)
tic()
res1 <- sdcol(A)
toc()

tic()
res2 <- apply(A,2,sd)
toc()

#install.packages("matrixStats")
library(matrixStats)
tic()
res3 <- colSds(A)
toc()
```

What function performed best? The function with best performance was *colSds()*
What could be the difference the other two? This is a compiled function besides that it avoids
setting environment variables and duplication of data.

Main message: if you have a compiled function available use it instead of writing your own code.



# 3.

Use the *Rcpp* package to write a C++ function that adds three float numbers:

```{r eval=TRUE}
library(Rcpp)

cppFunction('float sum_cpp(float a, float b, float c ) {
  float sumc = a + b + c;
  return sumc;
}')
sum_cpp
sum_cpp(2,4,6)
```

## 4. Exercise (Challenge)

The following functions **func1, func2** update the entry *i,j* of a matrix A (they do the same thing).
One of them is faster than the other. 

```{r eval=TRUE}
library(tictoc)
A <- matrix(runif(1000000),nrow=1000)

func1 <- function(a) {
  for(j in 1:ncol(a)) {
    for(i in 1:nrow(a)) {
      if (a[i,j] > 0.5 ) {
        a[i,j] <- 0.5
      }
    }
  }
  a
}
tic()
B <- func1(A)
toc()

func2 <- function(a) {
  for(j in 1:dim(a)[2]) {
    for(i in 1:dim(a)[1]) {
      if (a[i,j] > 0.5 ) {
        a[i,j] <- 0.5
      }
    }
  }
  a
}

tic()
C <- func2(A)
toc()

```

```{r eval=TRUE}
library(lineprof)
prof1 <- lineprof(func1(A))
prof1

prof2 <- lineprof(func2(A))
prof2
```

Look at the columns allocated, release, and duplicated data. Can you draw a conclusion why one 
of the functions is faster that the other?

Answer: The first function is allocating more memory and doing many more duplications than the
second function. Because of that it is slower than the second one. More profound reason dim() function is a primitive function and R doesn't need
to do unnecessary duplications
