---
title: "Solutions"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski"
date: "Feb., 2021"
output: html_document
---

## 1. Exercise

Use the following template by filling the *FIXME* strings with the corresponding
R code. In this exercise, you will use the **tictoc** package to time different
parts of a R code and save the results in a log file:

```{r eval=TRUE}
library("tictoc")

tic.clearlog()
tic("Total execution time")

tic("time: var init")
A <- matrix(1.0, 5000, 5000)
toc(log = TRUE)

sumcol <- function(B) {

tic("time: var init func")
    l <- ncol(B)  #obtain the number of columns
  colsm <- rep(0,l) #create a vector to save the sums
toc(log = TRUE)

tic("time: for loop")
  for(j in 1:l){  
    s <- 0
    for(i in 1:l){
      s <- s + B[i,j]
    }
    colsm[j] <- s
  }
toc(log = TRUE)
  return(colsm)
}

res1 <- sumcol(A)

toc(log = TRUE)
logfile <- tic.log()
logfile
```

What is the most expensive part of this code? The loop

## 2. Exercise

In this problem you will use common packages to profile R code. Replace the *FIXME* strings
with appropriate R code.

Given the matrix A of ones with a size of 5000x5000:

```{r eval=TRUE}
A <- matrix(1.0, 5000, 5000)
```

compare the profiling results of the following functions in a) and b).

a) the user function *sumcol* computes the sum of the elements by columns


```{r eval=TRUE}
sumcol <- function(B) {
   l <- ncol(B)  #obtain the number of columns
   colsm <- rep(0,l) #create a vector to save the sums
   for(j in 1:l){  
      s <- 0
      for(i in 1:l){
          s <- s + B[i,j]
      }
      colsm[j] <- s
   }
   return(colsm)
}

Rprof("Rprofa.out")          #Save the Rprof output in a filename called Rprofa.out
res1 <- sumcol(A)
Rprof(NULL)

summaryRprof("Rprofa.out")   #view the profiling's summary 
```


b) the R built-in *colSums* function for computing the sums of elements by columns

```{r eval=TRUE}
Rprof("Rprofb.out")
res2 <- colSums(A)
Rprof(NULL)
summaryRprof("Rprofb.out") 
```


* Are the performances of the two functions similar? No
* The two functions do the same calculation, why the performaces could differ? The user-defined
function *sumcol* performs plenty of calls to R functions behind scenes.

## 3. Exercise

**Challenge:** Do a benchmarking of the previous two functions by using rbenchmark
and microbenchmark packages:

```{r eval=TRUE}
#initialize the matrix A and set function sumcol
A <- matrix(1.0, 5000, 5000)
sumcol <- function(B) {
   l <- ncol(B)  #obtain the number of columns
   colsm <- rep(0,l) #create a vector to save the sums
   for(j in 1:l){  
      s <- 0
      for(i in 1:l){
          s <- s + B[i,j]
      }
      colsm[j] <- s
   }
   return(colsm)
}

library(rbenchmark)
res3 <- benchmark(sumcol(A), colSums(A), replications=10)
res3 

library(microbenchmark)
res4 <- microbenchmark(sumcol(A), colSums(A), times=10)
res4
```
