---
title: "Solutions"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski, Lars Viklund"
date: "Feb., 2021"
output: html_document
---


## 2. 
 
* The following *foreach* loop divides each element in a row by the rows mean. 
Analyze the timings with *tictoc* for 1 and 2 cores:


```{r eval=TRUE}
library(doParallel)
library(tictoc)
A <- matrix(1.0, 5000, 2000)
cl <- makeCluster(1)  
registerDoParallel(cl)

tic()
res1 <- foreach(i=1:nrow(A), .combine='rbind') %dopar% (A[i,]/mean(A[i,])) 
toc()

stopCluster(cl)
```



* What are the timings when one goes from 1 core to 2 cores? Does the performance improve considerably upon increasing the number of cores? How could that be explained?

The performance doesn't increase considerably because of memory duplication.

## 3. 

* Obtain the timings of the *max.eig()* function, which computes the eigenvalues of a 
matrix of *NxN*, by using a different number of cores and using the **foreach** function. *N* can go from 1 to 700.

* How is the performance of the **foreach** function with 1 core compared with that of the **sapply** function (N=700) alternative? 

**foreach** function performs better, **sapply** function incurs in some extra overhead.


```{r eval=TRUE}
library(doParallel)
library(tictoc)

#Adapted from: https://rstudio-pubs-static.s3.amazonaws.com/181635_8d565543c7ce4218b80345adaa11ea58.html
max.eig <- function(N, sigma) {
    d <- matrix(rnorm(N**2, sd = sigma), nrow = N)
    E <- eigen(d)$values
    abs(E)[[1]]
}

cl <- makeCluster(1)
registerDoParallel(cl)
tic()
res2 <- foreach(n = 1:700, .combine='rbind') %dopar% max.eig(n, 1)
toc()
stopCluster(cl)
```


```{r eval=TRUE}
tic()
res3 <- sapply(1:700, function(n) {max.eig(n, 1)})
toc()
```
