---
title: "Solutions"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski"
date: "Feb., 2021"
output: html_document
---

## 0. Exercise

Modify the **job.sh** script in this folder by requesting 2 cores for running the
**HelloWorld.R** example. Although this R script uses 1 core (serial code), sometimes
it is useful to request more than 1. When do you think that this trick could be
helpful?

## 1. Exercise

Take a look at the code for the credit bank data set in the script **parallel.R** and also
at the **job_parallel.sh** batch script. Set the number of cores you want to use in both
scripts first to 1 and then increase the number up to 12 and get timings. 

Upload the working folder to Kebnekaise and run the simulation. Monitor the timings for 
the simulations. You may use the *job-usage* tool on Kebnekaise. Replace the *FIXME* 
strings with appropriate code.

## 2. Exercise
 
* The following *foreach* loop divides each element in a row by the rows mean. 
Analyze the timings with *tictoc* for 1 and 2 cores:


```{r eval=TRUE}
library(doParallel)
library(tictoc)
A <- matrix(1.0, 5000, 2000)
cl <- makeCluster(1)  
registerDoParallel(cl)

tic()
res1 <- foreach(i=1:nrow(A), .combine='rbind') %dopar% (A[i,]/mean(A[i,])) 
toc()

stopCluster(cl)
```

* Prepare and R script with the corrected code above (*FIXME* strings). And then 
modify the *job.sh* batch script to submit it to the SLURM queue. 

* Obtain the timings for 1 core to 2 cores. Does the performance improve considerably 
upon increasing the number of cores? How could that be explained? (**Hint: data
duplication**)

* What are the timings when one goes from 1 core to 2 cores? Does the performance improve considerably upon increasing the number of cores? How could that be explained?

The performance doesn't increase considerably because of memory duplication.

## 3. Exercise 

* Obtain the timings of the *max.eig()* function, which computes the eigenvalues of a 
matrix of *NxN*, by using a different number of cores and using the **foreach** function. 
*N* can go from 1 to 700. Replace *FIXME* strings:


You may copy/paste this code into a file (called maybe script0.R) in your account 
on Kebnekaise. Then, use the *job.sh* batch script  (with the proper modifications) 
to submit your job to the queue with *sbatch job.sh*:

```{r eval=TRUE}
library(doParallel)
library(tictoc)

#Adapted from: https://rstudio-pubs-static.s3.amazonaws.com/181635_8d565543c7ce4218b80345adaa11ea58.html
max.eig <- function(N, sigma) {
    d <- matrix(rnorm(N**2, sd = sigma), nrow = N)
    E <- eigen(d)$values
    abs(E)[[1]]
}

cl <- makeCluster(1)
registerDoParallel(cl)
tic()
res2 <- foreach(n = 1:700, .combine='rbind') %dopar% max.eig(n, 1)
toc()
stopCluster(cl)
```


* How is the performance of the **foreach** function with 1 core compared with that of the **sapply** function (N=700) alternative:

```{r eval=TRUE}
tic()
res3 <- sapply(1:700, function(n) {max.eig(n, 1)})
toc()
```


## 4. Exercise (Challenge)

This exercise was inspired by the Stackoverflow case 
(https://stackoverflow.com/questions/31575585/shared-memory-in-parallel-foreach-in-r)
where it was observed that memory increased when using **foreach** in parallel mode.
Prepare a R script with the following code which uses 2 cores (call the script
script1.R for instance):


```{r eval=TRUE}
library(parallel)
library(foreach)
library(bigmemory)
library(doParallel)

#create a matrix of size 1GB aproximatelly
c<-matrix(runif(10000^2),10000,10000)
#convert it to bigmatrix
x<-as.big.matrix(c)
# get a description of the matrix
mdesc <- describe(x)

p <- detectCores(logical = FALSE) # only physical cores
cl <- makeCluster(2)
registerDoParallel(cl)
clusterEvalQ(cl, library(bigmemory))
out<-foreach(linID = 1:10, .combine=c) %dopar% {
  m <- attach.big.matrix(mdesc)
  #dummy expression to test data aquisition
  Sys.sleep(190)
  print(summary(m))
}
stopCluster(cl)
```

Now, use the **job.sh** template in this folder to submit your script to the queue.
Don't forget to replace the project name and the reservation queue. 
Use the job_ID (number you get when submitting the job with **sbatch job.sh**)
to get a URL with  the command **job-usage job_ID**. Copy/paste this URL to 
your local browser and wait ~5min. to see the statistics. Don't close your browser.

Repeat the process above and create a second R script (maybe called script2.R) with
the code below.

```{r eval=TRUE}
library(parallel)
library(foreach)
library(bigmemory)
library(doParallel)

#create a matrix of size 1GB aproximatelly
c<-matrix(runif(10000^2),10000,10000)
#convert it to bigmatrix
x<-as.big.matrix(c)
# get a description of the matrix
mdesc <- describe(x)

p <- detectCores(logical = FALSE) # only physical cores
cl <- makeCluster(2)
registerDoParallel(cl)
clusterEvalQ(cl, library(bigmemory))
out<-foreach(linID = 1:10, .combine=c) %dopar% {
  m <- attach.big.matrix(mdesc)
  #dummy expression to test data aquisition
  c[1,1] <- 0.3
  Sys.sleep(190)
  print(summary(m))
}
stopCluster(cl)
```

Modify the **job.sh** batch script, submit it to the queue. Then, get a URL
with **job-usage job_ID** tool and visualize the output in your local machine.

Compare the graphical outputs of both simulations (on your web browser).
What are the differences? What could be the source of that?

