---
title: "Exercises"
author: "Pedro Ojeda, Birgitte Brydsö, Mirko Myllykoski, Lars Viklund"
date: "Feb., 2021"
output: html_document
---

## 1. Exercise

Take a look at the code for the credit bank data set in the script **parallel.R** and also
at the **job_parallel.sh** batch script. Set the number of cores you want to use in both
scripts first to 1 and then increase the number up to 12 and get some timings. 

Upload the working folder to Kebnekaise and run the simulation. Monitor the timings for 
the simulations. You may use the *job-usage* tool on Kebnekaise. Replace the *FIXME* 
strings with appropriate code.

## 2. Exercise
 
* The following *foreach* loop divides each element in a row by the rows mean. 
Analyze the timings with *tictoc* for 1 and 2 cores:


```{r eval=TRUE}
library(doParallel)
library(tictoc)
A <- matrix(1.0, 5000, 2000)
cl <- makeCluster(*FIXME*)        
registerDoParallel(cl)

tic()
res1 <- foreach(i=1:nrow(A), .combine='rbind') %dopar% (A[i,]/mean(A[i,])) 
toc()

stopCluster(cl)
```



* What are the timings when one goes from 1 core to 2 cores? Does the performance improve considerably upon increasing the number of cores? How could that be explained?

## 3. Exercise

* Obtain the timings of the *max.eig()* function, which computes the eigenvalues of a 
matrix of *NxN*, by using a different number of cores and using the **foreach** function. *N* can go from 1 to 700. Replace *FIXME* strings:

* How is the performance of the **foreach** function with 1 core compared with that of the **sapply** function (N=700) alternative?

```{r eval=TRUE}
library(doParallel)
library(tictoc)

#Adapted from: https://rstudio-pubs-static.s3.amazonaws.com/181635_8d565543c7ce4218b80345adaa11ea58.html
max.eig <- function(N, sigma) {
    d <- matrix(rnorm(N**2, sd = sigma), nrow = N)
    E <- eigen(d)$values
    abs(E)[[1]]
}

cl <- makeCluster(*FIXME*)
registerDoParallel(cl)
tic()
res2 <- foreach(n = 1:700, .combine='rbind') %dopar% max.eig(n, 1)
toc()
stopCluster(cl)
```

```{r eval=TRUE}
tic()
res3 <- sapply(1:700, function(n) {max.eig(n, 1)})
toc()
```
